The case for open preprints in biology

### 2013-03-20 ms.tex
master branch

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex ms
% bibtex ms
% latex ms
% latex ms
% dvipdf ms

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
\usepackage{setspace}
\usepackage{hyperref} 
\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
%\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{The Case for Open Preprints in Biology}
}
% Insert Author names, affiliations and corresponding author email.
\\
Philippe Desjardins-Proulx$^{1, 2,\ast}$, 
Ethan P. White$^{3}$, 
Joel J. Adamson$^{4}$,
Karthik Ram$^{5}$,
Timoth\'ee Poisot$^{1,2,6}$,
Dominique Gravel$^{1,2}$
\\
\bf{1} Theoretical Ecosystem Ecology laboratory, Universit\'e du Qu\'ebec \`a Rimouski, Quebec, Canada.
\\
\bf{2} Quebec Center for Biodiversity Science, McGill University, Quebec, Canada.
\\
\bf{3} Departement of Biology, Utah State University, Utah, United-States of America.
\\
\bf{4} Ecology, Evolution and Organismic Biology, University of North Carolina at Chapel Hill, North Carolina, United-States of America.
\\
\bf{5} Environmental Science, Policy, and Management, University of California, Berkeley, California, United-States of America.
\\
\bf{6} International Network for Next-Generation Ecology.
\\
$\ast$ E-mail: philippe.d.proulx@gmail.com
\end{flushleft}

\section*{Introduction}

@@

Despite the success of this approach in other fields, most manuscripts in
biology are not posted to preprint servers and are therefore not seen by more
than a handful of other scientists prior to publication. In this article, we
highlight the advantages of open preprint servers for both scientists and
publishers, discuss the preprint policies of major publishers in biology, and
describe the main options to publish preprints (Box 1, Table 2).

\section*{The case for public preprints}

The first and most often discussed advantage of open preprints is
speed (Figure 1). The time between submission and the official
publication of a manuscript can be measured in months, sometimes in years. For
all this time, the research is known only to a select few: colleagues, editors,
reviewers. Thus, the science cannot be used, discussed, or reviewed by the wider
scientific community. In a recent blog post, C. Titus Brown noted how posting a
paper on arXiv quickly led to a citation (arXiv papers can be cited) and his
research was used by another researcher \cite{bro12}. The current system of
hiding manuscripts before acceptance pose problems for both scientists and
publishers. Manuscripts that are unknown cannot be used and thus take more time
to be cited. It has been shown that high-energy physics, with its high arXiv
submission rate, had the highest immediacy among physics and mathematics
\cite{pra05}. Immediacy measures how quickly articles are cited.

@@

\section*{Preprints in biological sciences}

In contrast to other disciplines, the field of biology has effectively no
preprint culture, with the exception of small pockets of primarily highly
quantitative research (\emph{e.g.}, epidemiology, population genetics).  While
submitting to preprint servers has become more common in the past few years, the
number of biology papers submitted to preprint servers still represents only a
small fraction of the total research produced in biology (Figure 2).

@@

\section*{Box 1: Preprint Server Roundup}

\subsection*{arXiv}

@@

\subsection*{figshare}

@@

\subsection*{PeerJ}

@@

\subsection*{F1000Research}\label{f1000r}

@@

\subsection*{GitHub}

@@

\subsection*{Other options}

@@

\section*{Conclusion}

@@

\section*{Acknowledgements}

@@

\section*{References}
\bibliography{refs}

\newpage
\section*{Figure Legends}

\subsection{Figure 1 (map.tiff)}

It can take several months before a submitted paper is
officially published and citable.} Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are
given access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analyzed, and integrated
into current research.

\subsection{Figure 2 (arxiv.tiff)}

Submissions to the quantitative biology section lag behind
physics, mathematics, and computer science.

\newpage
\section*{Tables}

\begin{table}[!ht]
  \caption{\bf{Policies for important publishers in biology.}}
  \begin{tabular}{|ll|}
  \hline
  Publisher                             & Policy \\
  \hline
  Springer                              & Accept \\
  BMC                                   & Accept \\
  Elsevier                              & Accept \\
  Nature Publishing Group               & Accept \\
  Public Library of Science             & Accept \\
  Genetics Society of America           & Accept \\
  Royal Society                         & Accept \\
  National Academy of Science (USA)     & Accept \\
  Ecological Society of America         & Accept \\
  Oxford Journals                       & Accept \\
  Science                               & Ambiguous \\
  Wiley-Blackwell                       & No general policy \\
  British Ecological Society            & No answer to our query \\
  \hline
  \end{tabular}
  \label{table:policies}
\end{table}

\begin{table}[!ht]
  \caption{\bf{Popular options for preprints}}
  \begin{tabular}{|l|ccccccc|}
  \hline
  Website               & Free & Comments & Private & Peer-reviewed & DOI & Version-Control & Other Content\\
  \hline
  arXiv.org             & Yes & No & No & No & No & No & No\\
  figshare.com          & Yes & Yes & No & No & Yes & No & Yes\\
  peerj.com             & 1/yr & Yes & Yes & No & Yes & No & No\\
  f1000research.com     & No & Yes & No & Yes & Yes & No & No\\
  github.com            & Yes & Yes & Yes & No & No & Yes & Yes\\
  \hline
  \end{tabular}
  \label{table:box1}
\end{table}

\newpage
\section*{Table Legends}

\subsection*{Table 1}

Some publishers tolerate preprints except for a few of their medical
journals, for example the \emph{ Journal of the National Cancer Institute}
from Oxford and \emph{The Lancet} from Elsevier.

\subsection*{Table 2}

\textbf{Free:} Can preprints be submitted for free.
\textbf{Comments:} Support for online comments.
\textbf{Private:} Support for private preprints.
\textbf{Peer-reviewed:} Whether the preprints are peer-reviewed on the server.
\textbf{DOI:} Each item is assigned a unique digital object identifier.
\textbf{Version-Control:} Is the preprint stored using a version-control system with the complete history of modifications?
\textbf{Other content:} Can upload figures, videos, datasets, code.

>>>>

\documentclass[letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage{color,graphicx,hyperref,authblk}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}
\renewcommand\Affilfont{\itshape\small}
\begin{document}

\title{The case for open preprints in biology}

\author[1,2,3]{Philippe Desjardins-Proulx}
\author[4]{Ethan P. White}
\author[5]{Joel J. Adamson}
\author[6]{Karthik Ram}
\author[2,3,7]{Timoth\'ee Poisot}
\author[2,3]{Dominique Gravel}
\affil[1]{email: \href{philippe.d.proulx@gmail.com}{philippe.d.proulx@gmail.com}}
\affil[2]{Quebec Center for Biodiversity Science, McGill University, Montr\'eal, Canada.}
\affil[3]{Theoretical Ecosystem Ecology laboratory, Universit\'e du Qu\'ebec \`a Rimouski, Canada.}
\affil[4]{Departement of Biology, Utah State University, United-States of America.}
\affil[5]{Ecology, Evolution and Organismic Biology, University of North Carolina at Chapel Hill, United-States of America}
\affil[6]{Environmental Science, Policy, and Management. University of California, Berkeley, United-States of America.}
\affil[7]{International Network for Next-Generation Ecology.}

\maketitle

\section{Introduction}

@@

Despite the success of this approach in other fields, most manuscripts in
biology are not posted to preprint servers and are therefore not seen by more
than a handful of other scientists prior to publication. In this article, we
highlight the advantages of open preprint servers for both scientists and
publishers, discuss the preprint policies of major publishers in biology, and
describe the main options to publish preprints (Table \ref{table:options}).

\section{The case for public preprints}

\begin{figure}[ht!] \centering\includegraphics[width=0.90\textwidth]
{map.pdf} \caption {It can take several months before a submitted paper is
officially published and citable. Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are
given access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analyzed, and integrated
into current research.} \label{fig:map} \end{figure}

The first and most often discussed advantage of open preprints is
speed (Figure \ref{fig:map}). The time between submission and the official
publication of a manuscript can be measured in months, sometimes in years. For
all this time, the research is known only to a select few: colleagues, editors,
reviewers. Thus, the science cannot be used, discussed, or reviewed by the wider
scientific community. In a recent blog post, C. Titus Brown noted how posting a
paper on arXiv quickly led to a citation (arXiv papers can be cited) and his
research was used by another researcher \cite{bro12}. The current system of
hiding manuscripts before acceptance pose problems for both scientists and
publishers. Manuscripts that are unknown cannot be used and thus take more time
to be cited. It has been shown that high-energy physics, with its high arXiv
submission rate, had the highest immediacy among physics and mathematics
\cite{pra05}. Immediacy measures how quickly articles are cited.

@@

\section{Preprints in biological sciences}

\begin{figure}[ht!] \centering\includegraphics[width=0.90\textwidth]
{arxiv.pdf} \caption {Submissions to the quantitative biology section lag behind
physics, mathematics, and computer science.} \label{fig:arxiv} \end{figure}

In contrast to other disciplines, the field of biology has effectively no
preprint culture, with the exception of small pockets of primarily highly
quantitative research (\emph{e.g.}, epidemiology, population genetics).  While
submitting to preprint servers has become more common in the past few years, the
number of biology papers submitted to preprint servers still represents only a
small fraction of the total research produced in biology (Figure \ref{fig:arxiv}).

@@

\begin{table}[!ht]
  \caption{\bf{Policies for important publishers in biology.}}
  \begin{tabular}{|ll|}
  \hline
  Publisher                             & Policy \\
  \hline
  Springer                              & Accept \\
  BMC                                   & Accept \\
  Elsevier                              & Accept \\
  Nature Publishing Group               & Accept \\
  Public Library of Science             & Accept \\
  Genetics Society of America           & Accept \\
  Royal Society                         & Accept \\
  National Academy of Science (USA)     & Accept \\
  Ecological Society of America         & Accept \\
  Oxford Journals                       & Accept \\
  Science                               & Ambiguous \\
  Wiley-Blackwell                       & No general policy \\
  British Ecological Society            & No answer to our query \\
  \hline
  \end{tabular}
  \caption{Some publishers tolerate preprints except for a few of their medical
journals, for example the \emph{ Journal of the National Cancer Institute}
from Oxford and \emph{The Lancet} from Elsevier.}
  \label{table:policies}
\end{table}

\section{Preprint Server Roundup}

\begin{table}[!ht]
  \caption{\bf{Popular options for preprints}}
  \begin{tabular}{|l|ccccccc|}
  \hline
  Website               & Free & Comments & Private & P.-R. & DOI & V.C. & O.C.\\
  \hline
  arXiv.org             & Yes & No & No & No & No & No & No\\
  figshare.com          & Yes & Yes & No & No & Yes & No & Yes\\
  peerj.com             & 1/yr & Yes & Yes & No & Yes & No & No\\
  f1000research.com     & No & Yes & No & Yes & Yes & No & No\\
  github.com            & Yes & Yes & Yes & No & No & Yes & Yes\\
  \hline
  \end{tabular}
  \caption{
    \textbf{Free:} Can preprints be submitted for free.
    \textbf{Comments:} Support for online comments.
    \textbf{Private:} Support for private preprints.
    \textbf{P.-R.:} Whether the preprints are peer-reviewed on the server.
    \textbf{DOI:} Each item is assigned a unique digital object identifier.
    \textbf{V.C.:} Is the preprint stored using a version-control system with the complete history of modifications?
    \textbf{O.C.:} Can upload figures, videos, datasets, code.
  }
  \label{table:options}
\end{table}

\subsection{arXiv}

@@

\subsection{figshare}

@@

\subsection{PeerJ}

@@

\subsection{F1000Research}\label{f1000r}

@@

\subsection{GitHub}

@@

\subsection{Other options}

@@

\section{Conclusion}

@@

\section{Acknowledgements}

@@

\section{Funding}

PDP is supported by an Alexander Graham Bell scholarship from the National
Sciences and Engineering Council of Canada. EPW is supported by a CAREER Award
from the National Science Foundation (DEB-0953694). JJA is supported by NSF
DEB-0614166 and NSF DEB-0919018.  TP is supported by a FQRNT-MELS post-doctoral
scholarship and 25 cents found by a coffee machine.  KR is supported by NSF
DEB-1021553. DG is funded by a Discovery Grant from the National Sciences and
Engineering Council of Canada and by the Canada Research Chair program.

\bibliography{refs}
\bibliographystyle{plain}
